---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# `pbbo` -- Prior by Bayesian Optimisation

<!-- badges: start -->
[![R-CMD-check](https://github.com/hhau/pbbo/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/hhau/pbbo/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

`pbbo` uses information you supply about the prior predictive distribution to help you find a prior for the parameters in your Bayesian model.

## Installation

You can install the development version of pbbo from [GitHub](https://github.com/hhau/pbbbo) with:

```{r, eval = FALSE}
# install.packages("devtools")
devtools::install_github("hhau/pbbo")
```

## An example

Suppose the target distribution is `N(0, 0.5^2)`.
We define the target LCDF and function to draws samples from the target:

```{r}
library(pbbo)
suppressPackageStartupMessages(library(ParamHelpers))

target_lcdf <- function(x) {
  Rmpfr::pnorm(x, mean = 2, sd = 0.5, log.p = TRUE)
}

target_sampler <- function(n) {
  rnorm(n = n, mean = 2, sd = 0.5)
}
```
and define a the prior predictive distribution of a model (and its hyperparameters):

```{r}
prior_predictive_sampler <- function(n, lambda) {
  rnorm(n = n, mean = lambda["mu"], sd = lambda["sigma"])
}

param_set <- makeParamSet(
  makeNumericParam(id = "mu", default = 0.2, lower = -50, upper = 50),
  makeNumericParam(id = "sigma", lower = 0, upper = 20, default = 0.2)
)
```

We can then call `pbbo` to get optimal values of λ = (μ, σ)

```{r}
library(futile.logger)

flog.threshold(futile.logger::INFO, name = "pbbo")
flog.appender(
  appender.file("../../logs/pbbo-test.log"),
  "pbbo"
)

pbbo_res <- suppressWarnings(pbbo(
  target_lcdf = target_lcdf,
  target_sampler = target_sampler,
  prior_predictive_sampler = prior_predictive_sampler,
  param_set = param_set,
  n_crs2_iters = 300,
  n_internal_prior_draws = 5e3,
  n_internal_importance_draws = 1e3,
  importance_method = "uniform",
  bayes_opt_batches = 1,
  bayes_opt_iters_per_batch = 50,
  bayes_opt_design_points_per_batch = 40
))

opt_lambda <- pbbo_res %>%
  get_best_lambda()

print(opt_lambda)
```

We can compare the prior predictive distribution at the optima against the
target:

```{r compare}
suppressPackageStartupMessages(library(ggplot2))
suppressPackageStartupMessages(library(dplyr))

n_samples <- 2e3

plot_tbl <- tibble(
  x = c(
    prior_predictive_sampler(n_samples, opt_lambda),
    target_sampler(n_samples)
  ),
  grp = rep(c("optima", "target"), each = n_samples)
)

ggplot(plot_tbl) +
  geom_line(
    mapping = aes(x = x, colour = grp),
    stat = "density"
  ) +
  labs(colour = "Type")
```
